15-744 Project Proposal

### 2011-09-28 proposal/proposal.tex
Finished (tentatively) project proposal.

It's good for network applications to be able to support many simultaneous
connections---this is known as the C10K (``10,000 client'') problem\cite{c10k}.

Simultaneous client connections require an application to maintain multiple
states at once. A classic way to do this is to spawn a thread for each client.
However, this approach does not scale as the number of clients grows.

(nonblocking?)
(lightweight threads?)
(cite Erlang)

A lightweight method is to use \t{epoll} or \t{kquery}.
This requires the application programmer to manually maintain a state machine
with

Requires a lot of complex but straightforward code which is difficult to debug.
Unsurprisingly, there exist many thin C/C++ wrappers, but these still require
...

and provide no assurance that the program ...

@@

We present a domain-specific language which provides a simple

easily extensible with more protocol definitions

outputs C code to be linked with the rest of the server code.

formal guarantees: given our abstract specification of \t{epoll}, our DSL
generates code with the following properties...

embedded in Haskell: natively allows functional metaprogramming,
leverages native type system

@@

>>>>

Modern network applications need to support many simultaneous clients---this is
known as the C10K (``10,000 client'') problem\cite{Kegel}. But to maintain these
simultaneous connections, an application must track each client's state
independently. A straightforward way to do this is to spawn a thread for each
client, but this approach does not scale as the number of clients grows.

A lightweight approach is to use kernel mechanisms for multiplexed, non-blocking
I/O, such as \t{select}, \t{poll}, \t{epoll}, or \t{kquery}; unfortunately, to
maintain separate state for each client, application programmers must manually
structure their code as state machines, which is tedious and error-prone.
Unsurprisingly, there exist many thin C/C++ wrappers for these facilities, but
none avoid major code refactoring.

To maintain the illusion of threads (in particular, client-specific state),
languages like Erlang and Haskell provide lightweight threading via user-space
schedulers and non-blocking IO. While this affords better performance, the full
generality of this approach still imposes unnecessary overhead on networking
applications which don't require preemptable, independently scheduled threads
\cite{Vinoski}.

@@

Hence, we intend to produce an embedded domain-specific language in Haskell to
describe concurrent network applications. This language will provide composable,
high-level network operations, in which programmers may express their
application logic naturally, as if using threads.

However, this high-level description will be automatically synthesized into a
state machine representation usable in a C loop invoking \t{epoll}, using no
threads whatsoever. We will additionally output a graphical representation of
this underlying state machine.

The back-end of this system can be implemented via several different options
which we are still evaluating. One option is to directly interface with the
\t{epoll} hooks in the Haskell runtime system, bypassing the native lightweight
threads. Another option is to generate C source code which itself implements the
network application, optionally linkable with native C code.

@@

We will first collect a representative sample of Linux network applications and
examine the range of high-level primitives required to support them. We will
then decide on a specific back-end for our language, and design the front- and
back-end in tandem. To evaluate the utility of this language, we will build a
number of representative applications in our language (e.g., an echo server,
simple HTTP server) and compare them to traditional implementations, with
respect to both code simplicity and performance.
